Assessing Material Inventory Uncertainties in Integrated Fuel Cycle Simulations

### 2018-06-18 dis_clusion.tex
some work on the flows

It is really interesting to observe that when the cascade is left completely
untouched (Model A) or when it is slightly retuned to maintain the tail to
product enrichment factor as well as the cut of each centrifuges, chaining the
cascade we can observe high increase of the enrichment at each level.  On the
contrary, when retuning the cut of each centrifuges to maintain the ideal state
of the cascades while chaining them, the \gls{HEU} production rate is favored
over the enrichment gain.

The tail recycling allows for each model a huge gain in productivity, even for
then model B for which it does not change the number of levels required to reach
$90w\%$ of $^{235}$U in the uranium. Even if no cascades chaining also to
retrieve the same production rate as a direct enrichment, the model B with
reprocessing reach about $80\%$ of an optimum production, which is far from
being negligible\ldots

>>>>

It is really interesting to observe that when the cascade is left completely
untouched (Model A) or when it is slightly retuned to maintain the tail to
product enrichment factor as well as the cut of each centrifuges (Model C),
chaining the cascade we can observe high increase of the enrichment at each
level.  On the contrary, when retuning the cut of each centrifuges to maintain
the ideal state of the cascades (Model B) while chaining them, the \gls{HEU}
production rate is favored over the enrichment gain.

The tail recycling allows for each model a huge gain in productivity, even for
then model B where the number of levels required to reach $90w\%$ of $^{235}$U
in the uranium does not change. Even if no cascades chaining allows to retrieve
the same production rate as a direct enrichment, the model B with tail
reprocessing reach about $80\%$ of an optimum production, which is far from
being negligible, allowing to build up a Significant Quantity of \gls{HEU} in
less than 8 months\ldots

### 2018-06-18 experiement.tex
add ref +add SQ mention
some work on the flows

\section{the experiment}
This works focuses on comparing the different miss use models to a reference
calculation in which a single large cascade is build and designed to directly
produce \gls{HEU} from natural uranium.
This works uses the Cyclus fuel cycle simulator to allow material exchange
between facilities.

\subsection{The cascade configuration}
\subsubsection{reference}
As mentioned previously, all the further calculations will be compared to the
most favorable configuration to produce \gls{HEU}, where all the available
centrifuges are used in a single large cascade designed to directly produce \gls{HEU}
from natural uranium, with a tail assay close to $0.3w\%$ uranium. The design
characteristic of the reference cascade are summarized in Table
\ref{tab:cascade_config}.

\subsubsection{default cascade}
The default cascade is the cascade design for normal civilian enrichment
operation, enriching natural uranium to about $3.5w\%$. This cascade will be
layered and fed with uranium at higher enrichment to evaluate the possibility
to used them, with few or no tuning, to produce \gls{HEU}. The characteristics of
the default cascade are summarized in Table \ref{tab:cascade_config}.

@@

\subsection{scenarios}
Seven different simulations have been simulated:
\begin{itemize}
\item one as the reference calculation, with a single centrifuges designed to
    produce directly \gls{HEU} from natural uranium,
\item three calculations (one per miss-use model) were default cascade are
    chained to produce \gls{HEU}, without recycling the tails of each cascade,
    the cascade tails are directly sent to the waste,
\item three calculations (one per miss-use model) were default cascade are
    chained to produce \gls{HEU}, the tails of each cascade are recycled, blending
    the tail of one level in the feed of the previous level of
    cascades (see Figure \ref{fig:cascade_level}.
\end{itemize}
In the following, cascades can be connected in tandem, where each set of cascade
in parallel is called a ``level``, as illustrated on Figure
\ref{fig:cascade_level}.

\begin{figure}[ht] % replace 't' with 'b' to force it to be on the bottom
    \centering
    \includegraphics[scale=0.45]{flow}
    \caption{Schematic representation of the chained cascades with three levels,
    with the feed, product and the tail flows, respectively in blue, green and
    red. The dashed orange line represent the alternative tail flow with
    recycling them.}
    \label{fig:cascade_level}
\end{figure}

\subsection{Level population}
In order to assign the optimum number of cascade to each level a virtual cut as
been computed as:
\begin{equation}
    \theta^{v}_{i} = \frac{F_{i}-T_{i}}{P_{i}-T_{i}},
\end{equation}
where, i represents a level of cascade and $F_{i}$, $P_{i}$ and $T_{i}$
respectively the feed, product and tail assay of the cascades at this level.

@@

>>>>

\section{The experiment}
This works focuses on comparing the different miss-use models to a reference
calculation in which a single large cascade is build and designed to directly
produce \gls{HEU} from natural uranium.
This works uses the Cyclus fuel cycle simulator to allow material exchange
between facilities. The enrichment cascade algorithm have been implemented in
the mbmore package \cite{mbmore.2018}.
In each cases, 5060 centrifuges have been used and spread across up to 30
different gas enrichment cascades.

\subsection{The cascade configuration}

\subsubsection{reference}

As mentioned, all the further calculations will be compared to the most
favorable configuration to produce \gls{HEU}, where all the available
centrifuges are used in a single large cascade designed to directly produce
\gls{HEU} from natural uranium, with a tail assay close to $0.3w\%$. The design
characteristic of the reference cascade are summarized in Table
\ref{tab:cascade_config}.

\subsubsection{default cascade}

The default cascade is the cascade design for normal civilian enrichment
operation, enriching natural uranium to about $3.5w\%$, with a tail assay close
to $0.3w\%$. This cascade will be layered and fed with uranium at higher
enrichment to evaluate the possibility to used them, with few or no tuning, to
produce \gls{HEU}. The characteristics of the default cascade are summarized in
Table \ref{tab:cascade_config}.

@@

\subsection{scenarios}
In the following, cascades can be connected in tandem, where each set of cascade
in parallel is called a ``level``, as illustrated on Figure
\ref{fig:cascade_level}.
Seven different simulations have been simulated, to evaluate the effectiveness
of an enrichment cascade when used outside of its designed scope :
\begin{itemize}
\item one as the reference calculation, with a single centrifuges designed to
    directly produce \gls{HEU} from natural uranium,
\item three calculations (one per miss-use model) were default cascade are
    chained to produce \gls{HEU}, without recycling the tails of each cascade,
    the cascade tails are directly sent to the waste,
\item three calculations (one per miss-use model) were default cascade are
    chained to produce \gls{HEU}, and the tails of each cascade are recycled, blending
    the tail of one level in the feed of the previous level of
    cascades (see Figure \ref{fig:cascade_level}).
\end{itemize}

\begin{figure}[ht] % replace 't' with 'b' to force it to be on the bottom
    \centering
    \includegraphics[scale=0.45]{flow}
    \caption{Schematic representation of the chained cascades with three levels,
    with the feed, product and the tail flows, respectively in blue, green and
    red. The dashed orange line represent the alternative tail flow when
    tail recycling is considered.}
    \label{fig:cascade_level}
\end{figure}

\subsection{Level population}
In order to assign the optimum number of cascades to each level, a virtual cut as
been computed as:
\begin{equation}
    \theta^{v}_{j} = \frac{F_{j}-N''_{j}}{N'_{j}-N''_{j}},
\end{equation}
where, $j$ represents a level of cascade and $N_{j}$, $N'_{j}$ and $N''_{j}$
respectively the feed, product and tail assay of the cascades at this level.

@@

As it is not possible to assign a fraction of an enrichment cascade, cascade per
level are rounded up, to the first level is attibuted the remaining available
cascades.

\subsection{Timestep effect}

As it can be observed on Figures \ref{fig:case1}, \ref{fig:case2} and
\ref{fig:case3}, the different cascade levels required up to $2n+1$ timesteps
before it starts to enrich uranium, where $n$ corresponds to the level number.
This corresponds to the time required for the material to propagate from the
natural uranium source to the n-th level, regardless of the duration of the
timestep (hour, day, month,...). This is also true for the number of timesteps
required to reach the equilibrium assays value in the tail reprocessing cases.
As timestep can be artificially small or large, one will only be considering
equilibrium values flow rates and assays in the following analysis.

### 2018-06-18 motivation.tex
some work on the flows

The present work investigates the performance of a enrichment cycle when chaining
gas enrichment cascade tuned for low enrich uranium production from natural
uranium. As literature on the mater is for obvious reason limited, three
behavior models have been implemented and used to evaluate the response of an
enrich cascade when fed with different assays than the design one. This work
takes also advantage of the Cyclus\cite{cyclus} fuel cycle capabilities to
evaluate the assay blending equilibrium.

>>>>

The present work investigates the performance of an enrichment cycle when chaining
gas enrichment cascades tuned for low enrich uranium production from natural
uranium. As literature on the mater is for obvious reason limited, three
behavior models have been implemented and used to evaluate the response of an
enrich cascade when fed with different assays than the design one. This work
takes also advantage of the Cyclus\cite{cyclus} fuel cycle capabilities to
evaluate the assay blending equilibrium.

### 2018-06-18 results.tex
some work on the flows

As illustrated on Figures \ref{sfig:case1_NR}, \ref{sfig:case2_NR} and
\ref{sfig:case3_NR} and summarized on Tab \ref{tab:level}, the different model
don't have the same effect on the cascade behavior. While the models A and C,
allow a quick enrichment gain, with the cascades chaining, respectively
4/23/78/98 and 4/23/82/99, the model B, the enrichment gain is only
4/21/61/90...  The same effect is observed when recycling the tails\ldots

@@

As demonstrated on Figures \ref{sfig:case1_R}, \ref{sfig:case2_R} and
\ref{sfig:case3_R}, recycling the tails increases the overall product assay at
all the different levels. As shown on Table \ref{tab:level}, the tail assay of
level $n$ is always higher than the product assay of level $n-2$, recycling the
tails of level $n$ will consequently increase the feed assay of level $n-1$.
Moreover, as the feed assay of level $n-1$ increases, its tail and product
assays increase as well, increasing de facto the feed assays of respectively
cascade levels $n-2$ and $n$...  This effect allow to reduce the number of
cascade levels required to reach \gls{HEU} in case A and C.

@@

As Model A and C, relies on maintaining the cut values at each stages of the
cascade and share the same number of levels, both the  same repartition of
cascades across the different levels and then the same \gls{HEU} production
rate.

>>>>

As illustrated on Figures \ref{sfig:case1_NR}, \ref{sfig:case2_NR} and
\ref{sfig:case3_NR} and summarized on Tab \ref{tab:level}, the different model
don't have the same effect on the cascade behavior. While the models A and C,
allow a quick enrichment gain, with the cascades chaining, respectively
4/23/78/98 and 4/23/82/99, the model B, the enrichment gain is only
4/21/61/90.  Similar effects can be observed when recycling the tails\ldots

@@

As shown on Figures \ref{sfig:case1_R}, \ref{sfig:case2_R} and
\ref{sfig:case3_R}, recycling the tails increases the overall product assay at
all the different levels. As the tail assay of a level $n$ is always higher than
the product assay of the level $n-2$, recycling the tails of level $n$ will
consequently increase the feed assay of level $n-1$ (see Table \ref{tab:level}).
Moreover, with an increased feed assay, tail and product assays increase as
well, increasing de facto the feed assays of respectively cascade levels $n-2$
and $n$...  This effect reduces the number of cascade levels required to reach
\gls{HEU} in case A and C.

@@

As Model A and C, relies on maintaining the cut values at each stages of the
cascade and share the same number of levels, have the exact same cascade
repartition across the different levels and the same \gls{HEU} production rate.

### 2018-06-18 theory.tex
discussion and conclusion
some work on the flows

\section{theory}
\subsection{Centrifuge properties}

The present work uses the analytical solution by R\"aetz \cite{raetz.phd} of the
differential equation for the gas centrifuge as described in \cite{glaser.2008}.
Centrifuge parameters, such as average gas temperature, $T$, peripheral speed,
$v$, height, $h$, diameter, $d$, pressure ratio, $x$, feed flow rate, $F$,
counter-current flow ratio, $L/P$, and efficiency, $e$ have been chosen (Table
\ref{tab:centrifuges}) to match the cascade design describe in
\cite{glaser.2008} and \cite{walker.2017}.  These parameters for a P1-type
centrifuge are used to estimate the JCPOA-compliant IR-1 centrifuge.

@@

The cascade is built as an ideal cascade, with no losses in the separative work,
which corresponds to $\alpha =\beta = const$ for all stage of the cascade, where
$\alpha$ and $\beta$ respectively represent the feed to product and the feed to
tail enrichment factors.  $\alpha$ and $\beta$ can be expressed as function of
the abundance ($R$) or the enrichment ($N$) of respectively the product
($R'$,$N'$) and the feed,($R$,$N$) and the feed and the tails ($R''$,$N''$) such
as:
\begin{subequations} \label{eqs:alphabeta}
    \begin{equation} \label{eq:alpha_def}
        \alpha = \frac{R'}{R} = \frac{N'}{1-N'}\frac{1-N}{N} 
\end{equation}
\\
\begin{equation}\label{eq:beta-def}
        \beta = \frac{R}{R''} = \frac{N}{1-N}\frac{1-N''}{N''} 
\end{equation}
\end{subequations}

As detailed in \cite{avery} it is also possible to derive $\alpha$ from the
first principle, and express it as a function of the feed rate $F$ the
separative performance $\delta U(\theta)$, the cut $\theta$:

@@

Since $\alpha_{i}$ and $\beta_{i}$ remain constant, only the value of the cut,
$\theta_{i}$, changes in each stage $i$ of a cascade.  This algorithm assumes
that the corresponding separative power $\delta U$ (not re-computed) can be
achieved with the chosen centrifuge design, tuning other operational parameter
such as the rotation speed, the counter-current flow ratio...  Once $\theta_{i}$
is determined, it is possible to compute the product and the tail assay.

The design of the cascade is performed through 2 steps.  First one determines the
configuration and number of stages, adding stages until the product assay of the
final stage is greater than or equal to than the desired assay, and the tails
assay is similarly less than or equal to the desired tails assay.  This
determines the number of enriching and stripping stages as well as their
enrichment properties ($N_{i}$, $N'_{i}$, $N''_{i}$,$\theta_{i}$i).

The second step determines how to populate the cascade with the user-defined
maximum number of centrifuges.

One solves the linear flow equation, \eqref{eq:flow}, to determine the
theoretical flow in the cascade.

@@

Once the relative flow of each stage has been determined, the cascade can be
populated with actual machines up the stages until the maximum number available
of machines is reached.

\subsection{Miss-use models}
Little information is available about optimising an existing enrichment cascade
that is being fed with a feed enrichment that does not match the design one. So
far 3 different methods have been investigated, the first one assumes that no
change are been made on the cascade, the second one, assumes that the cut value
at each stage is retuned to maintain the ideal state of the cascade, the last
one, described in \cite{walker.2017} assumes the tails to product enriching
factor remains constants ($\gamma = \alpha\times\beta$). Models behavior, and
assumptions are summarized in Tab. \ref{tab:models}.

@@

\subsubsection{Model A}
The tuning method A does not re-optimize $\theta_i$ based on the true flow
enrichment. As $\delta U$ and $\alpha$ do not depend on the stage feed assay
($N'$), they do not change from stage to stage. According to equation
\eqref{eq:beta}, when $\alpha$ and $\theta$ are fixed, when the feed assay ($N$)
changes, then $\beta$ will change accordingly.  This breaks the ideal status of
the cascade, i.e.  $N_{i} \neq N'_{i-1} \neq N''_{i+1}$.

In order to compute the proper product and tails assay at each stage, the
tails and the product from respectively the next and the previous stage must be
blended in order to determine the correct stage feed assay. As this is a obvious
cycling problem, an iterative solution has been chosen: all feed assays are
iteratively updated, blending the proper product and tails, then using the
updated feed assay, the new product and tail assays are recomputed. This process
is repeated until the change in assays is smaller than the set precision (1e-8
by default).
As the cut remain fixed at each stage the different flow do not need to be
recomputed.

@@

As the cuts values change, the flow rate between the different stage has to be
recomputed. Because the cascade is not reorganised (the number of cascade per
stage remain the same as the original design). The new flow rate are computed as
the flow rates of the reconfigured cascade scaled down by the possible
flows of the original one: the flow rates of a reconfigured cascade are
computed, the ratios of the flow stage by stage with the original are determined,
the smaller ratio is used to scale down all the flow rates.

@@

The only solution allowing product assay values ranging between 0 and 1 is the
following :
\begin{equation}\label{eq:model_b_sol}
    \frac{\gamma(N + \theta) - N 
            - \sqrt{\gamma^2(N^2 
                            - 2N\theta 
                            + \theta^2) 
                    - \gamma(2N^2 - 2\theta^2 
                            + 2N + 2\theta) 
                    + N^2 + 2N\theta + \theta^2 - 2N - 2\theta + 1} 
            - \theta + 1}
        {2\theta(\gamma - 1)}
\end{equation}
Once the product assay is known, one can trivially determine the tail assay,
$\alpha$ and $\beta$ using equations \eqref{eqs:alphabeta} and mass
conservation.

>>>>

\section{Theory}
\subsection{Centrifuge properties}

The present work uses the analytical solution by R\"aetz \cite{raetz.phd} of the
differential equation for the gas centrifuge as described in \cite{glaser.2008}.
Centrifuge parameters, such as average gas temperature, $T$, peripheral speed,
$v$, height, $h$, diameter, $d$, pressure ratio, $x$, feed flow rate, $F$,
counter-current flow ratio, $L/P$, and efficiency, $e$ have been chosen (Table
\ref{tab:centrifuges}) to match the cascade design describe in
\cite{glaser.2008} and \cite{walker.2017} using P1-type centrifuges.

@@

The cascade is built as an ideal cascade, with no losses in the separative work,
which corresponds to $\alpha =\beta = const$ for all stages of the cascade, where
$\alpha$ and $\beta$ respectively represent the feed to product and the feed to
tail enrichment factors.  $\alpha$ and $\beta$ can be expressed as function of
the abundance ($R$) or the enrichment ($N$) of respectively the product
($R'$, $N'$) and the feed ($R$, $N$) and the feed and the tails ($R''$, $N''$) such
as:
\begin{subequations} \label{eqs:alphabeta}
    \begin{equation} \label{eq:alpha_def}
        \alpha = \frac{R'}{R} = \frac{N'}{1-N'}\frac{1-N}{N} 
\end{equation}
\\
\begin{equation}\label{eq:beta-def}
        \beta = \frac{R}{R''} = \frac{N}{1-N}\frac{1-N''}{N''} 
\end{equation}
\end{subequations}

As detailed in \cite{avery} it is also possible to derive $\alpha$ from the
first principle, and express it as a function of the feed rate $F$, the
separative performance $\delta U(\theta)$ and the cut $\theta$:

@@

As $\alpha_{i}$ and $\beta_{i}$ remain constant, only the value of the cut,
$\theta_{i}$, changes across the different stages of a cascade.  This algorithm
assumes that the corresponding separative power $\delta U$ (not re-computed) can
be achieved with the chosen centrifuge design, tuning other operational
parameter such as the rotation speed, the counter-current flow ratio...  Once
$\theta_{i}$ is determined, it is possible to compute the product and the tail
assay.

The design of the cascade is performed through 2 steps.  First one determines
the configuration and number of stages, adding stages until the product assay of
the final stage is greater or equal the product targeted assay, and similarly
the tails assay is less or equal the tails desired assay.  This determines the
number of enriching and stripping stages as well as their enrichment properties
($N_{i}$, $N'_{i}$, $N''_{i}$,$\theta_{i}$i).

The second step determines the relative flows at each stages, solving the linear
flow equation, \eqref{eq:flow}.
The cascade can then be populated with actual machines until the maximum number
available of machines is reached.

@@

\subsection{Miss-use models}

Little information is available about optimising an existing enrichment cascade
that is being fed with a feed enrichment that does not match the design one. So
far 3 different methods have been investigated.

The first one assumes that no change are been made on the cascade, i.e $\delta
U$, $F$ and $\theta$ are fixed across all stages. The second one assumes the cut
value at each stage is retuned to maintain the ideal state of the cascade,
$\alpha$ and $\beta$ remain fixed. The last one, described in \cite{walker.2017}
assumes the tails to product enriching factor and the cut remain constants ($\gamma =
\alpha\times\beta$). Models behaviors and assumptions are summarized in Tab.
\ref{tab:models}.

@@

\subsubsection{Model A}

The tuning method A does not re-optimize $\theta_i$ keeping the same flow as the
ideal configuration. From equation \eqref{eq:alpha}, maintaining $\delta U$, $F$
and $\theta$ unchanged implies $\alpha$ remains unchanged as well. According to
equation \eqref{eq:beta}, when $\alpha$ and $\theta$ are fixed, if the feed
assay ($N$) changes, $\beta$ will change accordingly.  This breaks the
ideal status of the cascade, i.e. $N_{i} \neq N'_{i-1} \neq N''_{i+1}$.

In order to compute the proper product and tails assay at each stage, the tails
and the product from respectively the next and the previous stage must be
blended in order to determine the correct stage feed assay. As this is a obvious
cycling problem, an iterative solution has been chosen: all feed assays are
iteratively updated, blending the proper product and tails, then using the
updated feed assay, the new product and tail assays are recomputed. This process
is repeated until the sum of the square difference in assays is smaller than the
precision (1e-8).  As the cut remain fixed at each stage the different flow do
not need to be recomputed.

@@

As the cuts values change, the flow rate between the different stage has to be
recomputed. Because the cascade is not reorganised (the number of cascade per
stage remain the same as the original design). The new flow rate are computed as
the flow rates of the reconfigured cascade scaled down to match the allowed
flows of the original one: the flow rates at each stage of a miss-used cascade
have to be inferior or equal to the initial cascade design.

@@

The only solution allowing product assay to range between 0 and 1 is the
following :
\begin{equation}\label{eq:model_b_sol}
    \frac{\gamma(N + \theta) - N 
            - \sqrt{\gamma^2(N^2 
                            - 2N\theta 
                            + \theta^2) 
                    - \gamma(2N^2 - 2\theta^2 
                            + 2N + 2\theta) 
                    + N^2 + 2N\theta + \theta^2 - 2N - 2\theta + 1} 
            - \theta + 1}
        {2\theta(\gamma - 1)}
\end{equation}
Once the product assay is known, one can trivially determine the tail assay,
$\alpha$ and $\beta$ using equations \eqref{eqs:alphabeta} and mass
conservation.
